---
title: "superbowl"
author: "Valerie McIlvaine"
date: "7/13/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)

```


```{r upload files}
data_ad <- read_csv("~/STA 518/Final Project/project_proposal/Superbowl ad csv - Sheet1 (2).csv")
data_ad_time <- read_csv("~/STA 518/Final Project/project_proposal/Ad  Time _ Category - Sheet1 (1).csv")
data_ad_cost <- read_csv("~/STA 518/Final Project/project_proposal/Cost per Super bowl ad 2000-2020 - Sheet1.csv")
data <-read_csv("~/STA 518/Final Project/project_proposal/combined  - Sheet1 (2).csv")
```
```{r joining two data sets }

# merge data and data_ad_cost together 
total <- inner_join(data, data_ad_cost, by="year")
total
dim(total)


#chr to Numeric cost_per_minute and cost_per_thirty_seconds
total11<-
total%>% as.data.frame %>%
  mutate(cost_per_minute=str_remove_all(cost_per_minute,"[$,]"),
         cost_per_minute=as.numeric(cost_per_minute)) %>%

  mutate(cost_per_thirty_seconds=str_remove_all(cost_per_thirty_seconds,"[$,]"),
        cost_per_thirty_seconds=as.numeric(cost_per_thirty_seconds))%>%
  
  mutate(cost_per_second = cost_per_minute/60 )%>%
  
  mutate(cost_of_ad= cost_per_second*ad_time) %>%
  mutate(cost_of_ad_in_millions =cost_of_ad/1000000 )%>%
  mutate(cost_per_second_in_thousands =cost_per_second/100000)
  
total11

 # mutate(currency = scales::dollar(cost_per_second))
  
 
  

```
```{r }

#total1<- bind_cols(data_ad, data_ad_time %>% select(-year, -brand))



```


```{r exploratory bar charts}

# count of brands use of gender
ggplot(data=total, aes(x=category)) +
  geom_bar(stat= "count", position="dodge",color= "black", aes(fill=celebrity))+ 
   ggtitle("Bar char of Super Bowl Ad Categories and Celebrity Apperances From 2000-2020")


```
```{r line graph}


  ggplot(data=total11, aes(cost_per_second)) +
  geom_histogram(binwidth = 10)
  
  #line graph
  ggplot(data=total11, aes(x=year, y=cost_per_second_in_thousands)) +
  geom_line(linetype = "dashed", color="blue", size=1)+
    theme(legend.position="bottom")+
  geom_point(color="blue")+
  ggtitle("Cost Per Second of Super Bowl Ads from 2000 to 2020")+
    ylab("Cost (Hundreds of Thousands of Dollars)")
  
  
  
  #histogram of distribtuion of cost_of_ad by Year
  
total11 %>%
  group_by(year)

  ggplot(data=total11, aes(y=year, fill=year, color=year)) +
  geom_histogram(alpha=0.5, position="dodge", bins=10)+
       scale_color_brewer(palette="") + 
  theme_classic()+theme(legend.position="top")
 
 
  ggplot(data=total11, aes(x=year)) +
  geom_bar( stat="count", position="dodge",colour="black", aes(fill=diversity))
     
   

  
```
```{r porportions }


 ggplot(data=total11, aes(x= year, fill=use_sex))+
   geom_bar(stat= "count", position="fill")+
   ggtitle("Porportion of SuperBowl Ads that use Sex Appeal from 2000 to 2020")


```


```
```{r boxplot, reorder }

#reordering boxplots based on median cost_of_ad
ggplot(data=total11, aes(x= fct_reorder(category, cost_of_ad_in_millions), y=cost_of_ad_in_millions))+
geom_boxplot(fill="lightblue")+
ggtitle("Boxplots of Cost Per Ad by Category")+
xlab("Ad Category")+
ylab("Ad Cost in Millions")


#jitter 
ggplot(data=total11, aes(x=fct_reorder(category,cost_of_ad_in_millions), y= cost_of_ad_in_millions, colour = interaction(category ))) + 
      geom_point(alpha = 0.3, position = "jitter") +
      geom_boxplot(alpha = 0, colour = "black")+
         ggtitle("Boxplots of Cost Per Ad by Category")+
         xlab("Ad Category")+
         ylab("Cost in Millions")
         
#ad Length by year and category        
total11 %>%
  group_by(category) %>%
  ggplot(aes(x = year, y = ad_time, group = category, color= category)) +
 geom_point(alpha = 0.3)+
   ggtitle("Ad Length by Year and Ad Category")+
         ylab("Ad Length in Seconds")
        
```



```{r tables}



```





